---
title: "BMI Calculator"
author: "TWW"
output: ioslides_presentation
---

## Introduction

-   The BMI (Body Mass Index) Calculator is a Shiny application that allows users to calculate their BMI based on height, weight, gender, and age.
-   It provides information about the BMI category and associated health risks.

## How to Use

1.  Enter your height in centimeters.
2.  Enter your weight in kilograms.
3.  Select your gender.
4.  Enter your age.
5.  Click the "Calculate BMI" button.
6.  The application will display your calculated BMI, BMI category, and associated health risks.

## BMI Formula

-The BMI is calculated using the following formula:

```{r, echo=TRUE}
bmi <- function(weight, height) {
  weight / (height / 100)^2
}

```

```{r}
#The BMI categories are defined as follows:
getBmiCategory <- function(bmi) {
  if (bmi < 18.5) {
    return("Underweight")
  } else if (bmi >= 18.5 && bmi < 25) {
    return("Normal weight")
  } else if (bmi >= 25 && bmi < 30) {
    return("Overweight")
  } else {
    return("Obese")
  }
}
```

## Health Risk

The health risks associated with each BMI category are:

```{r}
getHealthRisks <- function(bmi) {
  if (bmi < 18.5) {
    return("Increased risk of nutritional deficiencies and osteoporosis.")
  } else if (bmi >= 18.5 && bmi < 25) {
    return("Lowest risk of chronic diseases.")
  } else if (bmi >= 25 && bmi < 30) {
    return("Increased risk of type 2 diabetes, hypertension, and cardiovascular diseases.")
  } else {
    return("High risk of type 2 diabetes, hypertension, cardiovascular diseases, and certain cancers.")
  }
}
```

## Example Calculation

Let's calculate the BMI and associated information for a person with the following characteristics:

```{r}
height <- 175
weight <- 75
gender <- "Male"
age <- 35

bmi_result <- bmi(weight, height)
bmi_category <- getBmiCategory(bmi_result)
health_risks <- getHealthRisks(bmi_result)

cat("BMI:", round(bmi_result, 2), "\n")
cat("BMI Category:", bmi_category, "\n")
cat("Health Risks:", health_risks)
```
